### 2012-03-02 ibis_workshop_2010/slides.tex

\documentclass[ignorenonframetext,compress]{beamer}

\mode<presentation>

\usepackage{times}
\usefonttheme{structurebold}

\usepackage[english]{babel}
\usepackage{tabularx}
\usepackage{url}
\usepackage{zed-csp}
\usepackage{subfigure}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{color}

\renewcommand*{\thesubfigure}{}

\usetheme{Madrid}

\title[PODD -- Ontology-driven Repository]{PODD: An Ontology-driven Data Repository for Collaborative Phenomics Research}

\author[Yuan-Fang Li]{
    \textbf{Yuan-Fang Li}\\
    \url{liyf@itee.uq.edu.au}
    }

\institute[UQ]{
    The eResearch Lab, School of ITEE\\The University of Queensland\\
    }
\vspace{+80pt}
\date[IBS Workshop 2010]{\footnotesize IBS Phenomics Data and Informatics Workshop - Canberra 2010}

\AtBeginSection[]{\frame{\frametitle{Outline}\tableofcontents[current]}}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\part{Main Part}

\section{Introduction}
\begin{frame}{Motivation}
\begin{block}{Challenges in phenomics research data management}
    \begin{itemize}
        \item Data is huge
        \begin{itemize}
            \item APN: estimated 1.8TB + 30\% growth/year
            \item TPA: 0.5TB/week $\sim 25$ TB/year
        \end{itemize}
        \pause
        \item New platforms/processes/technologies emerge
        \pause
        \item Data needs \textbf{\emph{context}}
        \begin{itemize}
            \item Scientific, administrative \& other metadata
        \end{itemize}
    \end{itemize}
\end{block}
\pause
\begin{block}{Repositories for the management of data}
    \begin{itemize}
      \item Not all questions answered
    \end{itemize}
\end{block}
\end{frame}

There are 3 major challenges to data management in phenomics research.

1. The volume of data. Phenomics research generates large amounts of
data. For example, APN, the Australian Phenomics Network, is
estimated to generate 1.8TB of data in the first year with 30\%
growth. That is more than 7TB over 3 years. And APN is only one
research center. We still have APPF, which operates 2 centers. The
large volume makes the storage and distribution much more difficult.

2. Phenomics is a new area. As a result, new imaging platforms,
technologies and processes emerge all the time. A data repository
needs to be constructed in such a way that it can cater for future
platforms and technologies.

3. Data needs context. 7TB of data isn't very useful if it just sits
there and collects dust. To make use of the data, we need to
surround them with context, the metadata, that describe the raw data
in a meaningful way. Metadata can have dimensions, such as
scientific and administrative. They can help researchers to discover
data, integrate data and analyze data is a vital part of a data
repository.

There are quite a few databases for genomics and protein research,
such as Emsembl and UniProt. Unfortunately, not all the requirements
for phenomics data management are met. For example, genomic data is
relatively well defined, which isn't the case for phenomics.
Phenomics data may come off from a variety of imaging and measuring
platforms and the appropriate metadata needs to be captured.

The above challenges motivated us to develop the PODD repository.

\begin{frame}{Clients \& Collaborators}
\begin{block}{Australian Integrated Biological Science Facilities}
    \begin{itemize}
        \item<2-> Australian Plant Phenomics Facility (APPF)
            \begin{itemize}
                \item High-throughput (TPA) \& high-resolution (HRPPC) centers
            \end{itemize}
        \item<3-> Australian Phenomics Network (APN)
            \begin{itemize}
                \item Mouse models, \emph{deep} imaging and measuring platforms
            \end{itemize}
        \item<4-> Atlas of Living Australia (ALA)
            \begin{itemize}
                \item Biodiversity data collection \& management
            \end{itemize}
    \end{itemize}
\end{block}
~\\
\begin{figure}[ht]
\centering
\onslide<2->\subfigure[]{
\includegraphics[width=20mm]{conveyor.png}
}
\onslide<2->\subfigure[]{
\includegraphics[width=20mm]{sample.png}
}
\onslide<3->\subfigure[]{
\includegraphics[width=20mm]{mouse.png}
}
\onslide<4->\subfigure[]{
\includegraphics[width=20mm]{organism.png}
}
\end{figure}
\end{frame}

The PODD repository is designed to be used by a number of Australian
research organizations and projects.

Australian Plant Phenomics Facility has two research centers.

The Plant Accelerator (TPA) at the University of Adelaide is a
high-throughput plant research center. What you see here is a
conveyor belt system. RFID-tagged plants travel in the system and
their images are automatically taken by the imaging stations.

The High Resolution Plant Phenomics Center, the HRPPC, is located
here in Canberra. The center focuses on deep phenotyping of plants
using a number of equipments. Some of them operate on single plants
in controlled environments, others operate on fields.

The Australian Phenomics Network (APN) is a collaborative research
facility established to provides mouse models for human and animal
diseases. For example, APN uses digital slide scanner to perform
histopathology and organ pathology imaging and analysis.

The Atlas of Living Australia is developing a biodiversity data
management system for Australia. Collaboration with ALA is also one
of PODD's goals.

\begin{frame}{Data Management Requirements}
\begin{overprint}
\onslide<1>
\begin{block}{Data capturing}
\begin{tabularx}{\linewidth}{p{35mm}X}
Flow Cytometry & FACS data \\
Histopathology & Zeiss slide images \\
Plant imaging & Lemnatec images, Flourogroscan images, 3D imaging \\
Infrared imaging & FLIR images \\
Chemical measurements & Chlorophyll content, Stomatal conductance \\
Visual observation & Manual reports\\
$\cdots$ & $\cdots$
\end{tabularx}
\end{block}

\onslide<2>
\begin{block}{Data generation}
\begin{tabularx}{\linewidth}{p{30mm}X}
Project & Project proposal, project plan \\
Investigation & Objectives, design \\
Materials & Lines/genotypes, samples, growth conditions \\
Devices & Specs, settings, versions \\
Processes & Workflows, protocols, variations \\
Measurements & Data, images \\
Analysis & Observations, results \\
$\cdots$ & $\cdots$
\end{tabularx}
\end{block}

\onslide<3>
\begin{block}{Data management}
\begin{tabularx}{\linewidth}{p{50mm}l}
Data distribution & ? \\
Data sharing & ? \\
Data publishing & ? \\
Access control & ? \\
Archival \& versioning & ? \\
Structure \& metadata & ? \\
Data discovery \& analysis & ?
\end{tabularx}
\end{block}
\end{overprint}
\end{frame}

One of the requirements of PODD is to support data capturing from
various imaging platforms.

As you can see on this slide. there are quite a number of different
ways of imaging and observing the subject. The challenge here is,
besides capturing the raw image files and observations, capture
sufficient metadata.

Another requirement is data generation. Data and metadata can be
generated on different levels and at different times. All this
information needs to be recorded in the repository.

Last but not least, there is a need for the effective management of
data. Tasks such as distribution, publishing, access control and
versioning belong to this category. As far as we know, there is no
data repository that handles these tasks for phenomics research.

\begin{frame}{PODD: an ontology-driven repository}
\begin{block}{Goals}
    \begin{itemize}
        \item Supports data captured by different platforms
        \item Supports different data formats
        \item Supports effective data management
        \begin{itemize}
            \item Metadata, distribution, access control, discovery, etc.
        \end{itemize}
    \end{itemize}
\end{block}
\pause
\begin{block}{Approach}
    \begin{itemize}
        \item Define domain ontologies -- focus of this talk
        \item Design an appropriate architecture
        \item Develop an ontology-aware repository
    \end{itemize}
\end{block}
\end{frame}

So, the goals of developing the PODD repository is to support the
above data management tasks for the phenomics community.

After our initial study, we realized that the traditional
database-driven approach may not be the best way of doing it. We
will come and revisit this point later.

Ontologies are a formal representation of a particular application
domain. Besides being a controlled vocabulary that define important
concepts and relations, an ontology can define complex constraints
on these concepts and relations, making it a very expressive
conceptualization of the domain.

Compared to database schemas, ontologies are more open and
extensible. For these reasons, we choose to develop PODD in an
ontology-based approach. In this talk I will focus on describing how
the domain ontology we develop is used in the repository in the
lifecycles of the management of domain objects.

\begin{frame}{PODD: The High-level Architecture}
 \vspace{-6mm}
 \begin{figure}[t]
  \begin{center}
  \includegraphics[trim=48mm 20mm 10mm 35mm, clip, width=130mm]{architecture.pdf}
   \end{center}
  \label{fig}
 \end{figure}
\end{frame}

Here you see the high-level architecture of the PODD repository. On
the basis are a few components. Fedora Commons is the underlying
repository that we use to store all the metadata information. iRODS
is a federated data storage software that can be used by Fedora
Commons. ANDS provides an iRODS-based service called Data Fabric,
which we will use in PODD to host the data captured.

Ontologies will be stored in Fedora Commons as well as in a RDF
triple store to enable data discovery using SPARQL queries and full
text search on metadata. We also use databases to store information
about users, access rights and some other things.

The data access layer works with ontologies and objects and is
responsible for serializing and deserializing objects from the
underlying storage.

The various services and the web interface is guarded by a security
layer that controls access based on fine-grained user access rights.

In the following few slides I will quickly introduce a few projects
related to PODD. I'll also briefly introduce the OWL ontology
language.

\section{Related Work}
\begin{frame}{Related Work}
\begin{block}{FuGe -- Functional Genomics Experiment}
    \begin{itemize}
        \item \emph{Material}, \emph{Protocol}, \emph{Data}, etc.
        \item Can be extended to support phenomics
        \item[$\cross$] May be difficult to extend for new concepts
    \end{itemize}
\end{block}
\pause
\begin{block}{OBI -- Ontology for Biomedical Investigations}
    \begin{itemize}
        \item \textit{``An integrated ontology for the description of life-science and clinical
        investigations.''}
        \item Comprehensive: 2,600+ classes, 10,000+ axioms
        \item[$\cross$] Complex, computationally ($\mathcal{SHOIN}(D)$)
    \end{itemize}
\end{block}
\end{frame}

FuGe is a UML and database based conceptual model for functional
genomics experiments. It models important domain concepts such as
material, protocol and data. The model is stored as database
schemas, as a result, FuGe may be difficult to extend.

The OBI ontology - ontology for biomedical investigation is a
comprehensive but complex ontology. It defines domain concepts 
starting from a very high abstract level. As a result, it is 
quite big, containing about 26 hundred classes and 10 thousand
axioms. It is also computationally expensive. For this reason and
that we need to performing reasoning repeatedly in PODD, we chose
not to use OBI as our domain ontology.

\begin{frame}{Related Work}
\begin{block}{Web Ontology Language (OWL)}
\begin{itemize}
    \item Precise, open \& extensible -- exactly what we need!
    \item Provides core vocabularies for expressing complex ontologies -- \emph{data models}
    \item Main language constructs
    \begin{itemize}
        \item Classes, e.g, \texttt{Human}
        \item Predicates , e.g., \texttt{hasParent}
        \item Individuals, e.g., \texttt{Aristotle}
    \end{itemize}
    \item APIs, query engines \& automated reasoners available
\end{itemize}
\end{block}
\end{frame}

The Web Ontology Language is the ontology language with a precise semantics 
and is designed to be open and extensible. For example, OBI is developed
in OWL. It provides language constructs to define domain concepts and relations
and axioms on them.

The three main ingredients in OWL are classes, predicates and individuals. 
For example, Human can be defined as a class; hasParent can be defined as
a predicate and Aristotle can be defined as an individual.

Quite a few open-source APIs, query engines and reasoners are available for OWL.

\section{The PODD Ontology}
\begin{frame}{The Ontology Approach}
\begin{block}{Benefits}
    \begin{itemize}
    \setlength{\itemsep}{6pt}
    \item<1-> Greater extensibility
        \begin{itemize}
        \item New concepts/relations can be easily added/modified
        \end{itemize}
    \item<2-> Better reuse \& integration
        \begin{itemize}
        \item Ontologies are \emph{open}
        \item Other ontologies can be integrated on multiple levels
        \end{itemize}
    \item<3-> Balance between expressivity \& reasoning complexity
        \begin{itemize}
        \item Formal semantics enables automated query \& analysis
        \item Off-the-shelf tools available
        \end{itemize}
    \end{itemize}
\end{block}
\end{frame}

The ontology-based approach has a number of benefits.

Firstly, the repository can be more extensible. In PODD we will store all
object information in ontologies. In such a way new concepts and relations
can be added more easily than if the information is stored in a database.

An ontology-based repository is also better in terms of reuse and integration.
Ontologies are designed to be shared so when you make your ontology available,
you share your model. When you publish your instance data, concrete domain data
is shared. So integration can happen on multiple levels.

The OWL language is defined with a formal semantics. This enables automated
query answering and analysis. OWL is designed with expressivity and reasoning
complexity in mind so that the core reasoning tasks can be made fully automated. 
As shown in a previous slide, although OWL is a relatively young language,
there are off-the-shelf tools available to use.

\begin{frame}{The PODD Ontology}
\begin{block}{Modeling essentials}
    \begin{itemize}
    \item Domain entities
        \begin{itemize}
        \item Abstract concepts
        \item Concrete objects
        \end{itemize}
    \item Domain entities defined using OWL \emph{ontologies}
    \end{itemize}
\end{block}
\pause
\begin{block}{Modeling in OWL}
    \begin{itemize}
    \item Domain concepts -- \emph{OWL classes}
    \item Inter-concept relations -- \emph{OWL predicates} \& \emph{OWL restrictions}
    \item Concrete domain objects -- \emph{OWL Individuals}
    \item Comments, descriptions -- \emph{OWL annotations}
    \end{itemize}
\end{block}
\end{frame}

In the same spirit of FuGe and OBI, the PODD ontology defines 
essential domain concepts so that we can describe concrete domain
objects. These concepts and objects are defined using OWL ontologies
and RDF triples.

\begin{frame}{The PODD Ontology -- An Example}
\begin{example}{The \textbf{\emph{Project}} concept}
    \begin{itemize}
    \item The top-level concept
    \item Constraints on inter-object relations \& attributes -- like everything else
    \end{itemize}
    \pause
    {\small
    \begin{align*}
    Project &\sqsubseteq~ =~ 1~ hasProjectPlan \sqcap \forall hasProjectPlan.ProjectPlan\\
            &\sqsubseteq~ \geq~ 1~ hasInvestigation \sqcap \forall hasInvestigation.Investigation\\
            &\sqsubseteq~ =~ 1~ hasStartDate \sqcap \forall hasStartDate.date\\
            &\sqsubseteq~ \leq~ 1~ hasPublicationDate \sqcap \forall hasPublicationDate.date
    \end{align*}}
\end{example}
\begin{block}{}
    \begin{itemize}
    \item \textcolor{blue}{\textbf{Extensibility}} from inheritance of OWL classes \& predicates
    \end{itemize}
\end{block}
\end{frame}

In this slide I show a simple, incomplete modeling of the top-level concept
in our ontology, the Project. In our model of phenomics, everything starts 
from a research project and all other objects, such as Investigation and
Material are child or descendant objects.

As you will see, we define constraints on attributes and inter-concept relations
of the Project concept. This is how we constrain all of our concepts.

Shown here are 8 constraints. The first line says that any Project instance must
be mapped to exactly 1 other object by the hasProjectPlan object and this 1 object
must be of type ProjectPlan.

\begin{frame}{The PODD Ontology}
\onslide<2->
\begin{columns}[t]
\column{.5\textwidth}
\begin{itemize}
    \item Inspired by FuGe \& OBI
    \item Aim: define all \emph{essential} domain concepts, attributes \& relations
    \item $\sim30$ classes, \\$\sim 80$ predicates, \\$\sim 200$ Axioms
    \item Faster reasoning, querying, etc.
\end{itemize}
\vfill
\onslide<1->
\column{.5\textwidth}
\vspace{-10mm}
 \begin{figure}[t]
  \begin{center}
  \includegraphics[trim=3mm 26mm 2mm 8mm, clip, height=72mm]{hier.pdf}
   \end{center}
 \end{figure}
\end{columns}
\end{frame}

Here you see a list of important domain concepts in the PODD ontology.
In the development of our ontology we took inspiration from FuGe and OBI
but with the aim that the PODD ontology will cover essential domain
concepts and relations without becoming too complex.

As it currently stands, the ontology defines about 30 OWL classes, 80
predicates and more than 200 logical axioms.

With this smaller ontology, all core reasoning tasks can be performed
much faster.

\begin{frame}{PODD Ontology -- Roles}
\begin{block}{Ontologies \emph{drives} repository functions}
\begin{list}{}{}
\item[\textcolor{blue}{\textbf{Presentation}}] Drives the rendering of the web pages
    \begin{itemize}
    \item Object creation, editing, display, etc.
    \end{itemize}
\item[\textcolor{blue}{\textbf{Management}}] Object (de)serialization to/from ontologies
\item[\textcolor{blue}{\textbf{Validation}}] Ontology reasoning performed
    \begin{itemize}
    \item Validation on object type, cardinality, etc.
    \end{itemize}
\item[\textcolor{blue}{\textbf{Discovery}}] Multiple ways of finding information
    \begin{itemize}
    \item Queries using SPARQL
    \item Searches using Lucene
    \end{itemize}
\end{list}
\end{block}
\end{frame}

The PODD ontology is at the center of the repository and it drives quite a lot
of repository functions. All the object metadata is saved as RDF triples
conforming to the ontology definitions. When a web page of creating, editing
or displaying an object is requested, the ontologies are used to determine
what gets displayed and in what way.

The domain objects are serialized into ontologies when they are saved and 
deserialized when they are loaded from the underlying storage.

Before we save and publish an object, the object's triples are validated against
the domain ontology. Objects can be queried and searched for on the RDF triples.

\begin{frame}{The PODD Ontology -- Object creation page rendering}
 \begin{figure}[t]
  \begin{center}
  \includegraphics[width=76mm]{create.png}
   \end{center}
 \end{figure}
\end{frame}

\section{Conclusion}
\subsection*{ }
\begin{frame}{Conclusion}
\begin{block}{To recap}
\begin{itemize}
\item Large amounts of data need to be managed
    \begin{itemize}
    \item There is a need for data archival, storage \& discovery
    \end{itemize}
\item Current approaches lacking/inadequate/inflexible
    \begin{itemize}
    \item Emerging processes, platforms, technologies require extensible infrastructure
    \end{itemize}
\item An ontology-driven approach as the foundation of PODD
    \begin{itemize}
    \item Extensible \& open
    \end{itemize}
\end{itemize}
\end{block}
\end{frame}

This slides shows a screen capture of the project creation page.
Note that most of these fields are generated by reading the domain
ontology and figuring out what fields are needed and what their types,
cardinalites are.

For example, in a previous slide we defined that a project must have 
exactly 1 start date attribute and it must be a date. So here we show 
an input box for this attribute and signal that it is mandatory. Note
that lead institution is also a mandatory attribute while the duration
is not. When this project object is saved, the values are checked against
the domain ontology and any errors or omissions will be reported back
to the user.

\begin{frame}{Conclusion}
\begin{block}{Where we are now}
\begin{itemize}
\item PODD ontology stabilizing
\item Development of basic repository functionality
\item Development of PODD web interface
\end{itemize}
\end{block}
\pause
\begin{block}{What's next}
\begin{itemize}
\item Development of batch data import/export processes
\item Development of object discovery services
\item Integrate with AAF federated authentication services
\item \textcolor{gray}{Exposing data for discovery}
\item \textcolor{gray}{Integrating with other data sources}
\end{itemize}
\end{block}
\end{frame}

\begin{frame}
	\frametitle{}
    \begin{center}
\Huge \textsc{\textcolor{blue}{\textbf{Thank You!}}}
\end{center}
\vspace{+20pt}
\begin{block}{Acknowledgment}
Faith Davies, Gavin Kennedy, Jane Hunter\\
eResearch Lab, Schoo of ITEE, UQ
\end{block}
\end{frame}

\end{document}